Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates

### 2014-10-29 response_to_reviewers.tex

\textbf{4. Response:}

{[}\textbf{\textgreater{}\textgreater{}\textgreater{}\textgreater{}
Marianne}, can you write whey we are not using the RC and PRC methods?
``The parametric rate comparison and relative rates test that he
suggests both require incomplete taxon sampling to be nonrandom, which
ours definitely isn't so we can easily explain why we didn't use
these.''\textbf{\textless{}\textless{}\textless{}\textless{}\textless{}\textless{}}{]}

The method that we used, BiSSE, might not be reliable when the number of
taxa is small, and there are biases in the number of species that have
each of the character states. However, we stated in the manuscript that
according to the evaluation of BiSSE by Davis, Midford, and Maddison
(2013), our data should not be affected for such biases.

The method BAMM can be used to explore the evolution of phenotypic
traits. However the trait characters
\href{http://bamm-project.org/configuration.html\#id5}{\textbf{must be
continuous}}. In our case the ``feeding on Solanaceae plants'' trait is
categorical (absence/presence) so it cannot be analyzed with BAMM.

BAMM also measures the diversification rates of particular clades. We
run an analysis to test whether we obtained congruent results. In the
Fig. 1, shown below, we can see that BAMM was able to pick up two
increases in diversification for Ithomiini butterflies, one for the
origin of the clade and, most importantly, another increase during the
recent diversification of the group. This is in agreement with our
results from using BiSSE, that the ithomiines using Solanaceae underwent
an increase in diversification rates.

\begin{figure}[htbp]
\centering
\includegraphics{response_to_reviewers_fig1.pdf}
\caption{Bamm analysis}
\end{figure}

\textbf{5. Reviewer's Comment:}

The authors should `play' with an AIC cutoff, simply to explore how it
might affect there inferences.

\textbf{5. Response:}

\begin{quote}
\color{blue}
As specified lines above and in the manuscript, we used the AIC cutoff
value 7.8 as suggested by MEDUSA (this value is regarded as optimal due
to the number of taxa in our analyses). Changing the cutoff value would
allow fewer or additional diversification shifts to be picked up by
MEDUSA. Hovewer the initial diversification shifts will remain the same
unless the AIC cutoff used is so hight that no single diversification
shifts are recovered. Thus playing with the AIC cutoff values does not
affect most of the diversification shifts discussed in our manuscript.
We rerun all our MEDUSA and MultiMEDUSA analysis with differente AIC
cutoff values to see any effect on the results. We used the AIC values 4
and 11.
\end{quote}

\begin{quote}
\color{blue}
The results in Fig.2 (also attached as PDF file) shows our MEDUSA
results from the manuscript (cutoff 7.8) along with results with AIC
cutoff values 4 and 11. We see that the first 19 diversification shitfs
are the same for cutoff values 7.8 and 4. However when using AIC cutoff
4 we recovered additional diversification shifts as the software kept
trying to fit additional models until the differente in AIC values was
less than 4. As expected MultiMEDUSA also found additinal
diversification shifts across the random sample of 1000 trees.
\end{quote}

\begin{quote}
\color{blue}
When we used AIC cutoff 11, we obtained fewer diversification shifts. We
recovered only 16, which are the same first 16 that we found in our
analysis with AIC 7.8.
\end{quote}

\begin{figure}[htbp]
\centering
\includegraphics{multimedusa_7_8_vs_4_vs_11_AIC_cutoff.pdf}
\caption{MEDUSA analyses under three AIC values}
\end{figure}

\textbf{6. Reviewer's Comment:}

In addition, the authors should be cautious about interpreting AIC in a
frequentist manner (e.g., ``Only three out of 13 significant shifts
found on the maximum credibility tree were consistent across more than
95\% of the trees from the posterior'') -- that is, AIC is a model
selection tool, which should act as a guide for selecting the `best'
family of models. Delta AIC should really not be thought of as alpha =
0.05.

\textbf{6. Response:}

\begin{quote}
\color{blue}
We agree that AIC values should not be treated in a frequentist manner.
However, we use the term \textbf{``significant shifts''}, in the same
way as the authors of the method MEDUSA (Alfaro et al. 2009), to
pinpoint to specific nodes that show a higher net diversification rate
than ancestral or descendant nodes.
\end{quote}

\begin{quote}
\color{blue}
We also used the AIC value in the correct way. We used the cutoff 7.8
for all MEDUSA analyses. Then we counted the number of significant
shifts infered in each tree and tried to pinpoint those that were
consistently recovered across the sample of 1000 random trees (and are
also present in the MCC tree). Since no single significant shift was
found in all 1000 trees we wanted to show those are were most commonly
recovered. This is way we considered consistent diversification shifts
those that were found in most of the trees (in more than 95\% of them).
\end{quote}

\textbf{7. Reviewer's Comment:}

Why are authors referred to as references, rather than by authors.

As noted by Reference 36, the responsible trait might\ldots{}

\textbf{7. Response:}

\begin{quote}
\color{blue}
We have fixed those references as suggested by the reviewer.
\end{quote}

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

Alfaro, Michael E, Francesco Santini, Chad Brock, Hugo Alamillo, Alex
Dornburg, Daniel L Rabosky, Giorgio Carnevale, and Luke J Harmon. 2009.
``Nine Exceptional Radiations Plus High Turnover Explain Species
Diversity in Jawed Vertebrates.'' \emph{P. Natl. Acad. Sci. USA} 106
(32): 13410--14.
doi:\href{http://dx.doi.org/10.1073/pnas.0811087106}{10.1073/pnas.0811087106}.

Davis, Matthew P, Peter E Midford, and Wayne Maddison. 2013. ``Exploring
power and parameter estimation of the BiSSE method for analyzing species
diversification.'' \emph{BMC Evol. Biol.} 13 (1): 38.
doi:\href{http://dx.doi.org/10.1186/1471-2148-13-38}{10.1186/1471-2148-13-38}.
